% \section{\thresam} \label{apx:thresh}
% For a discussion of the intuition behind the algorithm \thresam and 
% rigorous proof of
% Lemma~\ref{lemm:thresh}, we refer the reader to \shortciteS{Fahrbach2018}.

\section{\threshgreedy and Modification} \label{apx:threshgreedy}
\begin{algorithm}[t]
   \caption{The \threshgreedy Algorithm of \shortciteS{Badanidiyuru2014}}
   \label{alg:threshgreedy}
   \begin{algorithmic}[1]
     \Procedure{\threshgreedy}{$f, k, \epsi$}
     \State \textbf{Input:} evaluation oracle $f:2^{\mathcal N} \to \reals$, constraint $k$,
     accuracy parameter $\epsi > 0$
     \State $M \gets \argmax_{x \in \mathcal N} f(x)$; 
     \State $S \gets \emptyset$
     \For{ $\tau = M$; $\tau \ge (1 - \epsi)M/k$; $\tau \gets \tau (1 - \epsi)$}
     \For{ $x \in \mathcal N$ }
     \If{ $f( S + x ) - f(S) \ge \tau$ }
     \State $S \gets S + x$
     \If{ $|S| = k$ }
     \State \textbf{break} from outer \textbf{for}
     \EndIf
     \EndIf
     \EndFor
     \EndFor
     \State \textbf{return} $S$
     \EndProcedure
\end{algorithmic}
\end{algorithm}
In this section, we describe \threshgreedy (Alg.~\ref{alg:threshgreedy}) 
of \shortciteS{Badanidiyuru2014} and
how it is modified to have low adaptivity. This algorithm achieves 
ratio $1 - 1/e - \epsi$ in $O(n \log k)$ queries
if the function $f$ is monotone but has no constant ratio
if $f$ is not monotone. 

The \threshgreedy algorithm works as follows:
a set $S$ is initialized to the empty set. Elements
whose marginal gain exceed a threshold value are added
to the set in the following way:
initially,
a threshold of $\tau = M = \argmax_{a \in \mathcal N} f(a)$
is chosen, which is iteratively decreased by a factor
of $(1 - \epsi)$ until $\tau < M/k$. For each threshold $\tau$,
a pass through all elements
of $\mathcal N$ is made, during which any
element $x$ that satisfies $f(S + x) - f(S) \ge \tau$  is added to the set $S$.
While this strategy leads to an efficient $O(n \log k)$ total
number of queries, it also has $\Omega (n \log k)$ adaptivity, as
each query depends on the previous ones.

To make this approach less adaptive,
we replace the highly adaptive pass through $\mathcal N$ (the inner \textbf{for} loop)
with a single call to \thresam, which requires $O( \log n )$
adaptive rounds and $O(n / \epsi)$ queries in expectation.
This modified greedy approach appears twice in \latg (Alg.~\ref{alg:latg}),
corresponding to the two \textbf{for} loops. 
\section{Improved Ratio for \iter} \label{apx:iter}
\begin{algorithm}[t]
   \caption{The \iter Algorithm of \shortciteS{Gupta2010a}}
   \label{alg:iter}
   \begin{algorithmic}[1]
     \Procedure{\iter}{$f, k$}
     \State \textbf{Input:} evaluation oracle $f:2^{\mathcal N} \to \reals$, constraint $k$,
     \State $A \gets \emptyset$
     \For{ $i \gets 1$ to $k$ }
     \State $a_i \gets \argmax_{x \in \mathcal N} f( A + x ) - f(A)$
     \State $A \gets A + a_i$
     \EndFor
     \State $B \gets \emptyset$
     \For{ $i \gets 1$ to $k$ }
     \State $b_i \gets \argmax_{x \in \mathcal N \setminus A} f ( B + x ) - f (B)$
     \State $B \gets B + b_i$
     \EndFor
     \State $A' \gets $ \unc $(A)$
     \State \textbf{return} $C \gets \argmax \{ f(A), f(A'), f(B) \}$
     \EndProcedure
\end{algorithmic}
\end{algorithm}
In this section, we prove an improved approximation ratio
for the algorithm \iter of \shortciteS{Gupta2010a}, wherein a
ratio of $1/(4 + \alpha)$ is proven given access to
a $1/\alpha$-approximation for \unc. 
We improve this ratio to $\iterratio \approx 0.193$ if $\alpha = 2$.
Pseudocode for \iter
is given in Alg.~\ref{alg:iter}.

\iter works as follows. First a standard greedy procedure is run
which produces set $A$ of size $k$.
Next, a second greedy procedure is run to yield set $B$; during
this second procedure, elements of $A$ are ignored.
A subroutine for \unc is used on $f$ restricted to $A$, which
yields set $A'$. Finally the set of $\{A, A', B \}$ that maximizes
$f$ is returned. 
\begin{theorem}
  Suppose there exists an $(1/\alpha)$-approximation for
  \unc. Then by using this procedure as a subroutine,
  the algorithm \iter has approximation ratio
  $\iterratio$ for \sm.
\end{theorem}
\begin{proof}
  For $1 \le i \le k$, let $a_i,b_i$ be as chosen during the
  run of \iter. Define $A_i = \{ a_1, \ldots, a_{i - 1} \}$,
  $B_i = \{ b_1, \ldots, b_{i - 1} \}$. Then for 
  for any $1 \le i \le k$, we have
  \begin{align*}
    f(A_{i + 1}) + f(B_{i + 1}) - f(A_i) - f(B_i) &= f_{a_i}(A_i) + f_{b_i}( B_i ) \\
 &\ge \frac{1}{k} \sum_{o \in O} f_o( A_i ) + \frac{1}{k} \sum_{o \in O \setminus A} f_o (B_i ) \\
    &\ge \frac{1}{k} \left( \ff{O \cup A_i} - \ff{A_i} + \ff{ (O \setminus A) \cup B_i } - \ff{ B_i }\right) \\
    &\ge \frac{1}{k} \left( \ff{ O \setminus A } - (\ff{A_i} + \ff{B_i}) \right),
  \end{align*}
  where the first inequality follows from the greedy choices,
  the second follows from submodularity,
  and the third follows from submodularity and the fact that $A_i \cap B_i = \emptyset$.
  Hence, from this recurrence and standard arguments, 
  $$f(A) + f(B) \ge (1 - 1/e) \ff{O \setminus A},$$
  where $A,B$ have their values at termination of \iter.
  Since $f(A') \ge f(O \cap A) / \alpha$, 
  we have from submodularity
  \begin{align*}
    f(O) &\le f( O \cap A ) + f(O \setminus A) \\
    &\le \alpha f(A') + (1-1/e)^{-1}(f(A) + f(B)) \\
    &\le ( \alpha + 2(1-1/e)^{-1} ) f(C).
  \end{align*}
\end{proof}