
\documentclass[accepted,nohyperref]{article}

\usepackage{fullpage}
\usepackage{parskip}


\usepackage{authblk}


\usepackage{microtype}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{booktabs} %

\usepackage{hyperref}


\usepackage{xcolor}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}



\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{mathtools}
\usepackage{amsthm}

\usepackage[capitalize,noabbrev]{cleveref}

\theoremstyle{plain}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{assumption}[theorem]{Assumption}
\newtheorem{remark}[theorem]{Remark}


\definecolor{linkcolor}{RGB}{83,83,182}

\hypersetup{
    colorlinks=true,
    citecolor=linkcolor,
    linkcolor=linkcolor,
    urlcolor=linkcolor
}

\usepackage[backend=biber,natbib=true,style=authoryear,doi=false,isbn=false,url=false,eprint=false,giveninits=true,maxbibnames=200,maxcitenames=2,mincitenames=1,uniquename=false,uniquelist=false,dashed=false]{biblatex}

\newcommand{\noopsort}[1]{#1}


\addbibresource{references.bib}%

\DeclareSourcemap{
  \maps[datatype=bibtex]{
    \map{
      \step[fieldsource=url, match=\regexp{http://(dx.doi.org/|dl.acm.org/)}, final]
      \step[fieldset=url, null]
      \step[fieldset=urldate, null]
    }
  }
}
\setlength\bibitemsep{1ex}


\title{Differentially Private Coordinate Descent \\
  for Composite Empirical Risk Minimization}

\date{}

\author[1]{Paul Mangold}
\author[1]{Aurélien Bellet}
\author[2,3]{Joseph Salmon}
\author[4]{Marc Tommasi}

\affil[1]{Univ. Lille, Inria,  CNRS, Centrale Lille, UMR 9189 - CRIStAL, F-59000 Lille, France}
\affil[2]{IMAG, Univ Montpellier, CNRS, Montpellier, France}
\affil[3]{Institut Universitaire de France (IUF)}
\affil[4]{Univ. Lille, CNRS, Inria, Centrale Lille,  UMR 9189 - CRIStAL, F-59000 Lille, France}

\input{src/preamble}
\begin{document}

\maketitle



\begin{abstract}
  Machine learning models can leak information about the data used to
  train them. To mitigate this issue, Differentially Private (DP) variants of
  optimization
  algorithms like Stochastic Gradient Descent (DP-SGD) have been
  designed to trade-off utility for privacy in Empirical Risk
  Minimization (ERM) problems. In this paper, we propose
  Differentially Private proximal Coordinate Descent (DP-CD), a new
  method to solve composite DP-ERM problems. We derive utility
  guarantees through a novel theoretical analysis of inexact
  coordinate descent. Our results show that, thanks to larger step
  sizes, DP-CD can exploit imbalance in gradient coordinates to
  outperform DP-SGD. We also prove new lower bounds for composite
  DP-ERM under coordinate-wise regularity assumptions, that are nearly
  matched by DP-CD. For practical implementations, we propose to clip
  gradients using coordinate-wise thresholds that emerge
  from our theory, avoiding costly hyperparameter tuning. Experiments
  on real and synthetic data support our results, and show that DP-CD
  compares favorably with DP-SGD.
\end{abstract}

\input{src/intro}
\input{src/prelim}
\input{src/dp-cd}
\input{src/lower-bounds}
\input{src/practice}
\input{src/exp}
\input{src/related_work}
\input{src/conclu}

\section*{Acknowledgments}

The authors would like to thank the anonymous reviewers who provided
useful feedback on previous versions of this work, which helped to improve
the paper.

This work was supported in part by the Inria Exploratory Action FLAMED and
by the French National Research Agency (ANR) through grant ANR-20-CE23-0015
(Project PRIDE) and ANR-20-CHIA-0001-01 (Chaire IA CaMeLOt).


\printbibliography


\input{src/supplement}

\end{document}
