% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.21 of 2022/01/12
%
\pdfoutput=1
\RequirePackage{amsmath}
\documentclass[runningheads]{llncs}
%
\usepackage[T1]{fontenc}
% T1 fonts will be used to generate the final print and online PDFs,
% so please use T1 fonts in your manuscript whenever possible.
% Other font encondings may result in incorrect characters.
%

% other packages
\usepackage{amssymb}
\usepackage{multirow}
\usepackage[table,xcdraw]{xcolor}
\usepackage{textcomp} % for \textquotesingle
\usepackage{subfig}
\usepackage{algorithmic}
\usepackage{algorithm}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage[normalem]{ulem}
\usepackage{multirow}
\usepackage{booktabs}
\usepackage{threeparttable}
\usepackage{siunitx}
\usepackage[misc]{ifsym}
% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following two lines
% to display URLs in blue roman font according to Springer's eBook style:
%\usepackage{color}
%\renewcommand\UrlFont{\color{blue}\rmfamily}
%
\newcommand\doubleplus{+\kern-1.3ex+\kern0.8ex}
\begin{document}
%

\title{U-Net Inspired Transformer Architecture for \\Far Horizon Time Series Forecasting}

%
\titlerunning{Yformer for Far Horizon Time Series Forecasting}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
% \orcidID{0000-0001-6356-8646}
% \orcidID{1111-2222-3333-4444}
% \orcidID{0000-0002-7402-4166}
% \orcidID{2222--3333-4444-5555} 
% \orcidID{0000-0001-5729-6023}

\author{Kiran Madhusudhanan\inst{1}\Letter \and
Johannes Burchert \inst{1} \and
Nghia Duong-Trung \inst{2}\and
Stefan Born \inst{2} \and
Lars Schmidt-Thieme \inst{1} 
}
\authorrunning{K. Madhusudhanan et al.}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{Institute for Computer Science, University of Hildesheim, Hildesheim, Germany \\
\email{\{madhusudhanan, burchert, schmidt-thieme\}@ismll.uni-hildesheim.de}
\and
Technische Universit\"at Berlin, Berlin, Germany \\
\email{nghia.duong-trung@tu-berlin.de, born@math.tu-berlin.de}}

\toctitle{U-Net Inspired Transformer Architecture for Far Horizon Time Series Forecasting}
\tocauthor{Kiran~Madhusudhanan, Johannes~Burchert, Nghia~Duong-Trung, Stefan~Born and Lars~Schmidt-Thieme}

\maketitle              % typeset the header of the contribution
%
\begin{abstract}
Time series data is ubiquitous in research as well as in a wide variety of industrial applications. Effectively analyzing the available historical data and providing insights into the far future allows us to make effective decisions. Recent research has witnessed the superior performance of transformer-based architectures, especially in the regime of far horizon time series forecasting. However, the current state of the art sparse Transformer architectures fail to couple down- and upsampling procedures to produce outputs in a similar resolution as the input. We propose a U-Net inspired Transformer architecture named Yformer, based on a novel Y-shaped encoder-decoder architecture that (1) uses direct connection from the downscaled encoder layer to the corresponding upsampled decoder layer in a U-Net inspired architecture, (2) Combines the downscaling/upsampling with sparse attention to capture long-range effects, and (3) stabilizes the encoder-decoder stacks with the addition of an auxiliary reconstruction loss. Extensive experiments have been conducted with relevant baselines on three benchmark datasets, demonstrating an average improvement of 19.82, 18.41 percentage MSE and 13.62, 11.85 percentage MAE in comparison to the baselines for the univariate and the multivariate settings respectively.

\keywords{Time series Forecasting  \and Transformer \and U-Net}
\end{abstract}

\input{sections/introduction.tex}

\input{sections/relatedwork.tex}

\input{sections/background.tex}

\input{sections/methodology.tex}

\input{sections/experiments.tex}

\input{sections/ablation.tex}

\input{sections/conclusion.tex}


% % ---- Bibliography ----

% % BibTeX users should specify bibliography style 'splncs04'.
% % References will then be sorted and formatted in the correct style.

\bibliographystyle{splncs04}
\begin{thebibliography}{10}
    \providecommand{\url}[1]{\texttt{#1}}
    \providecommand{\urlprefix}{URL }
    \providecommand{\doi}[1]{https://doi.org/#1}
    
    \bibitem{layernorm}
    Ba, L.J., Kiros, J.R., Hinton, G.E.: Layer normalization. CoRR  (2016)
    
    \bibitem{BoxArima}
    Box, G.E.P., Jenkins, G.M.: Some recent advances in forecasting and control.
      Journal of the Royal Statistical Society. Series C (Applied Statistics)
      (1968)
    
    \bibitem{camacho2013model}
    Camacho, E.F., Alba, C.B.: Model predictive control. Springer science \&
      business media (2013)
    
    \bibitem{cao2018brits}
    Cao, W., Wang, D., Li, J., Zhou, H., Li, L., Li, Y.: Brits: Bidirectional
      recurrent imputation for time series. In: NeurIPS (2018)
    
    \bibitem{cirstea2022triformer}
    Cirstea, R.G., Guo, C., Yang, B., Kieu, T., Dong, X., Pan, S.: Triformer:
      Triangular, variable-specific attentions for long sequence multivariate time
      series forecasting. IJCAI  (2022)
    
    \bibitem{elu}
    Clevert, D., Unterthiner, T., Hochreiter, S.: Fast and accurate deep network
      learning by exponential linear units (elus). In: ICLR (2016)
    
    \bibitem{10.1145/3292500.3330662}
    Fan, C., Zhang, Y., Pan, Y., Li, X., Zhang, C., Yuan, R., Wu, D., Wang, W.,
      Pei, J., Huang, H.: Multi-horizon time series forecasting with temporal
      attention learning. In: SIGKDD (2019)
    
    \bibitem{he2016deep}
    He, K., Zhang, X., Ren, S., Sun, J.: Deep residual learning for image
      recognition. In: CVPR (2016)
    
    \bibitem{huang2017densely}
    Huang, G., Liu, Z., Van Der~Maaten, L., Weinberger, K.Q.: Densely connected
      convolutional networks. In: CVPR (2017)
    
    \bibitem{hyndman2018forecasting}
    Hyndman, R.J., Athanasopoulos, G.: Forecasting: principles and practice. OTexts
      (2018)
    
    \bibitem{Jarrett2020Target-Embedding}
    Jarrett, D., van~der Schaar, M.: Target-embedding autoencoders for supervised
      representation learning. In: ICLR (2020)
    
    \bibitem{jawed2019multi}
    Jawed, S., Rashed, A., Schmidt-Thieme, L.: Multi-step forecasting via
      multi-task learning. In: IEEE Big Data (2019)
    
    \bibitem{tungbcc19}
    Kieu, T., Yang, B., Guo, C., S.~Jensen, C.: Outlier detection for time series
      with recurrent autoencoder ensembles. In: IJCAI (2019)
    
    \bibitem{kitaev2020reformer}
    Kitaev, N., Kaiser, L., Levskaya, A.: Reformer: The efficient transformer. In:
      ICLR (2020)
    
    \bibitem{alexnet}
    Krizhevsky, A., Sutskever, I., Hinton, G.E.: Imagenet classification with deep
      convolutional neural networks. NeurIPS  (2012)
    
    \bibitem{lai2018modeling}
    Lai, G., Chang, W.C., Yang, Y., Liu, H.: Modeling long-and short-term temporal
      patterns with deep neural networks. In: SIGIR (2018)
    
    \bibitem{le2018supervised}
    Le, L., Patterson, A., White, M.: Supervised autoencoders: Improving
      generalization performance with unsupervised regularizers. NeurIPS  (2018)
    
    \bibitem{li2019enhancing}
    Li, S., Jin, X., Xuan, Y., Zhou, X., Chen, W., Wang, Y.X., Yan, X.: Enhancing
      the locality and breaking the memory bottleneck of transformer on time series
      forecasting. NeurIPS  (2019)
    
    \bibitem{tft}
    Lim, B., Arık, S.{\"O}., Loeff, N., Pfister, T.: Temporal fusion transformers
      for interpretable multi-horizon time series forecasting. Int. J. Forecast.
      (2021)
    
    \bibitem{lin2017feature}
    Lin, T.Y., Doll{\'a}r, P., Girshick, R., He, K., Hariharan, B., Belongie, S.:
      Feature pyramid networks for object detection. In: CVPR (2017)
    
    \bibitem{nbeats}
    Oreshkin, B.N., Carpov, D., Chapados, N., Bengio, Y.: N-beats: Neural basis
      expansion analysis for interpretable time series forecasting. In: ICLR (2020)
    
    \bibitem{perslev2019u}
    Perslev, M., Jensen, M., Darkner, S., Jennum, P.J., Igel, C.: U-time: A fully
      convolutional network for time series segmentation applied to sleep staging.
      NeurIPS  (2019)
    
    \bibitem{petit2021u}
    Petit, O., Thome, N., Rambour, C., Themyr, L., Collins, T., Soler, L.: U-net
      transformer: Self and cross attention for medical image segmentation. In:
      International Workshop on MLMI (2021)
    
    \bibitem{qin2017dual}
    Qin, Y., Song, D., Chen, H., Cheng, W., Jiang, G., Cottrell, G.W.: A dual-stage
      attention-based recurrent neural network for time series prediction. In:
      IJCAI (2017)
    
    \bibitem{RangapuramDeepState}
    Rangapuram, S.S., Seeger, M.W., Gasthaus, J., Stella, L., Wang, Y.,
      Januschowski, T.: Deep state space models for time series forecasting. In:
      NeurIPS (2018)
    
    \bibitem{ronneberger2015u}
    Ronneberger, O., Fischer, P., Brox, T.: U-net: Convolutional networks for
      biomedical image segmentation. In: MICCAI (2015)
    
    \bibitem{sagheer2019time}
    Sagheer, A., Kotb, M.: Time series forecasting of petroleum production using
      deep lstm recurrent networks. Neurocomputing  (2019)
    
    \bibitem{salinas2020deepar}
    Salinas, D., Flunkert, V., Gasthaus, J., Januschowski, T.: Deepar:
      Probabilistic forecasting with autoregressive recurrent networks. Int. J.
      Forecast.  (2020)
    
    \bibitem{sezer2020financial}
    Sezer, O.B., Gudelek, M.U., Ozbayoglu, A.M.: Financial time series forecasting
      with deep learning: A systematic literature review: 2005--2019. Applied soft
      computing  (2020)
    
    \bibitem{vaswani2017attention}
    Vaswani, A., Shazeer, N., Parmar, N., Uszkoreit, J., Jones, L., Gomez, A.N.,
      Kaiser, {\L}., Polosukhin, I.: Attention is all you need. In: NeurIPS (2017)
    
    \bibitem{wang2020linformer}
    Wang, S., Li, B.Z., Khabsa, M., Fang, H., Ma, H.: Linformer: Self-attention
      with linear complexity. ArXiv  (2020)
    
    \bibitem{wang2019srn}
    Wang, Y., Tao, X., Shen, X., Jia, J.: Wide-context semantic image
      extrapolation. In: CVPR (2019)
    
    \bibitem{wang2017time}
    Wang, Z., Yan, W., Oates, T.: Time series classification from scratch with deep
      neural networks: A strong baseline. In: IJCNN (2017)
    
    \bibitem{wu2021autoformer}
    Wu, H., Xu, J., Wang, J., Long, M.: Autoformer: Decomposition transformers with
      auto-correlation for long-term series forecasting. NeurIPS  (2021)
    
    \bibitem{zhou2020informer}
    Zhou, H., Zhang, S., Peng, J., Zhang, S., Li, J., Xiong, H., Zhang, W.:
      Informer: Beyond efficient transformer for long sequence time-series
      forecasting. In: AAAI (2021)
    
    \bibitem{zhou2021nnformer}
    Zhou, H.Y., Guo, J., Zhang, Y., Yu, L., Wang, L., Yu, Y.: nnformer: Interleaved
      transformer for volumetric segmentation. ArXiv  (2021)
    
    \bibitem{ZOU201939}
    Zou, X., Wang, Z., Li, Q., Sheng, W.: Integration of residual network and
      convolutional neural network along with various activation functions and
      global pooling for time series classification. Neurocomputing  (2019)
    
    \end{thebibliography}
    


\clearpage
\input{sections/appendix.tex}

    

\end{document}
