\section{Identification}\label{sec:id}

In seminal work, \cite{rosenbaum1983central,robins1986new} state sufficient conditions under which causal functions, philosophical quantities defined in terms of potential outcomes $\{Y^{(d)}\}$, can be measured from empirical quantities such as outcomes $Y$, treatments $D$, and covariates $(V,X)$. Colloquially, this collection of sufficient conditions is known as selection on observables. We assume selection on observables in the main text, and Pearl's front and back door criteria in Supplement~\ref{sec:graphical}.

\begin{assumption}[Selection on observables]\label{assumption:selection}
Assume
\begin{enumerate}
    \item No interference: if $D=d$ then $Y=Y^{(d)}$.
    \item Conditional exchangeability: $\{Y^{(d)}\}\indep D \mid X$.
    \item Overlap: if $f(x)>0$ then $f(d \mid x)>0$, where $f(x)$ and $f(d \mid x)$ are densities. 
\end{enumerate}
For $\theta_0^{CATE}$, replace $X$ with $(V,X)$.
\end{assumption}

No interference is also called the stable unit treatment value assumption. It rules out network effects, also called spillovers. Conditional exchangeability states that conditional on covariates $X$, treatment assignment is as good as random. Overlap ensures that there is no covariate stratum $X=x$ such that treatment has a restricted support.
%, similar to \cite{imbens2009identification}. %Figure~\ref{dag:te} presents a representative DAG \cite{pearl1993comment}. 
 To handle $\theta_0^{DS}$, we place a standard assumption in transfer learning.
\begin{assumption}[Distribution shift]\label{assumption:covariate}
Assume
\begin{enumerate}
    \item $\tilde{\text{\normalfont pr}}(Y,D,X)=\text{\normalfont pr}(Y \mid D,X)\tilde{\text{\normalfont pr}}(D,X)$;
    \item $\tilde{\text{\normalfont pr}}(D,X)$ is absolutely continuous with respect to $\text{\normalfont pr}(D,X)$.
\end{enumerate}
\end{assumption}
Populations $\text{\normalfont pr}$ and $\tilde{\text{\normalfont pr}}$ differ only in the distribution of treatments and covariates. Moreover, the support of $\text{\normalfont pr}$ contains the support of $\tilde{\text{\normalfont pr}}$.  An immediate consequence is that the regression $\gamma_0(d,x)=E(Y \mid D=d,X=x)$ remains the same across the different populations $\text{\normalfont pr}$ and $\tilde{\text{\normalfont pr}}$. 
